MS Technical Paper: \\ Placement Algorithms for Heterogeneous FPGAs

### 2025-09-11 latex/paper/sections/03-introduction.tex
basic edits

Field-Programmable Gate Arrays (FPGAs) have witnessed rapid growth in capacity and versatility, driving significant advances in computer-aided design (CAD) and electronic design automation (EDA) methodologies. 
Since the early-to-mid 2000s, the stagnation of single-processor performance relative to the rapid increase in integrated circuit sizes has led to a design productivity gap, where the computational effort for designing complex chips continues to rise. 
FPGA CAD flows mainly encompass synthesis, placement, and routing; all of which are NP-hard problems, of which placement is one of the most time-consuming processes. 
Inefficienct placement strategy not only extends design times from hours to days, thereby elevating cost and reducing engineering productivity, but also limits the broader adoption of FPGAs by software engineers who expect compile times akin to those of software compilers like {\tt gcc}.

For these reasons, FPGA placement remains a critical research effort even today. 
In this paper, we study and implement established placement methods. 
To do this, we use the RapidWright API, which is a semi-open-source research effort from AMD/Xilinx that enables custom solutions to FPGA design implementations and design tools that are not offered by their industry-standard FPGA environment, Vivado. 
We implement multiple variations of simulated annealing placers for Xilinx's 7-series FPGAs, with an emphasis on minimizing total wirelength while mitigating runtime. 
Our implementation is organized into three consecutive substages. 
The \textbf{prepacking} stage involves traversing a raw EDIF netlist to identify recurring cell patterns—such as CARRY chains, DSP cascades, and LUT-FF pairs—that are critical for efficient mapping and legalization. 
In the subsequent \textbf{packing} stage, these identified patterns, along with any remaining loose cells, are consolidated into \texttt{SiteInst} objects that encapsulate the FPGA’s discrete resource constraints and architectural nuances. 
Finally, the \textbf{placement} stage employs a simulated annealing (SA) algorithm to optimally assign \texttt{SiteInst} objects to physical \texttt{Sites}, aiming to minimize total wirelength while respecting the constraints of the 7-series architecture.

>>>>

Field-Programmable Gate Arrays (FPGAs) have witnessed rapid growth in capacity and versatility, driving significant advances in computer-aided design (CAD) and electronic design automation (EDA) methodologies. 
Since the early-to-mid 2000s, the stagnation of single-processor performance relative to the rapid increase in integrated circuit sizes has led to a design productivity gap, where the computational effort for designing complex chips continues to rise. 
FPGA CAD flows mainly encompass synthesis, placement, and routing; all of which are NP-hard problems, of which placement is one of the most time-consuming processes. 
Inefficient placement strategy not only extends design times from hours to days, thereby elevating cost and reducing engineering productivity, but also limits the broader adoption of FPGAs by software engineers who expect compile times akin to those of software compilers like {\tt gcc}.

For these reasons, FPGA placement remains a critical research effort even today. 
In this paper, we study and implement established placement methods. 
To do this, we use the RapidWright API, which is a semi-open-source research effort from AMD/Xilinx that enables custom solutions to FPGA design implementations and design tools that are not offered by their industry-standard FPGA environment, Vivado. 
We implement multiple variations of simulated annealing placers for Xilinx's 7-series FPGAs, with an emphasis on minimizing total wirelength while mitigating runtime. 
Our implementation is organized into three consecutive substages. 
The \textbf{prepacking} stage involves traversing a raw EDIF netlist to identify recurring cell patterns—such as CARRY chains, DSP cascades, and LUT-FF pairs—that are critical for efficient mapping and legalization. 
In the subsequent \textbf{packing} stage, these identified patterns, along with any remaining loose cells, are consolidated into \texttt{SiteInst} objects that encapsulate the FPGA’s discrete resource constraints and architectural nuances. 
Finally, the \textbf{placement} stage employs a simulated annealing (SA) algorithm to optimally assign \texttt{SiteInst} objects to physical \texttt{Sites}, aiming to minimize total wirelength while the constraints of the 7-series architecture.

### 2025-09-11 latex/paper/sections/12-future_work.tex
basic edits

with $i, j \in \{1, 2, ..., M\}$, where $M$ is the total number of modules in the design.
Module placements in the $x$ and $y$ directions are captured by the placement vectors \( \boldsymbol{x} = (x_1, x_2, ..., x_n) \) and \( \boldsymbol{y} = (y_1, y_2, ..., y_n) \), where $x_n$ and $y_n$ correspond to the coordinates of the $n$th module instance.
Edges connecting pins $i$ and $j$ have weights $w_{ij}$ and are assigned values depending on the selected net model and objective.

@@

The connectivity matrix $\boldsymbol{Q_x}$ encodes 2-pin connections between movable instances while $\boldsymbol{c_x}$ encodes 2-pin connections between movable instances and fixed instances   
Movable instances are for example \texttt{SLICEL}s, DSP, and RAM, while fixed instances can be IO pads and clock buffers.
We distinguish fixed instances because their placement cannot be optimized and can be thought of as constraints encoded into the objective.
For a design with $M$ pins belonging to movable instances and $F$ pins belonging to fixed instances, $\boldsymbol{Q}_x \in {\mathbb{R}}^{M \times M}$ and $\boldsymbol{c}_x \in {\mathbb{R}}^F$.

>>>>

with $i, j \in \{1, 2, ..., N\}$, where $N$ is the total number of modules in the design.
Module placements in the $x$ and $y$ directions are captured by the placement vectors \( \boldsymbol{x} = (x_1, x_2, ..., x_n) \) and \( \boldsymbol{y} = (y_1, y_2, ..., y_n) \), where $x_n$ and $y_n$ correspond to the coordinates of the $n$th module instance.
Edges connecting pins $i$ and $j$ have weights $w_{ij}$ and are assigned values depending on the selected net model and objective.

@@

The connectivity matrix $\boldsymbol{Q_x}$ encodes 2-pin connections between movable instances while $\boldsymbol{c_x}$ encodes 2-pin connections between movable instances and fixed instances   
Movable instances are for example \texttt{SLICEL}s, DSP, and RAM, while fixed instances can be IO pads and clock buffers.
We distinguish fixed instances because their placement cannot be optimized and can be thought of as constraints encoded into the objective.
For a design with $M$ pins belonging to movable instances and $F$ pins belonging to fixed instances, $\boldsymbol{Q}_x \in {\mathbb{R}}^{M \times M}$ and $\boldsymbol{c}_x \in {\mathbb{R}}^F$.
We can construct $\boldsymbol{Q}_x$ and $\boldsymbol{c}_x$ formally by taking the graph Laplacian matrix of our undirected weighted graph (with M-M and M-F connections included), then partitioning the matrix to separate M-M from M-F weights to form $\boldsymbol{Q}_x$ and $\boldsymbol{c}_x$.
Alternatively, we can simply follow a shortcut outlined in SimPL \cite{SimPL}, where we construct $Q_x$ by setting the diagonal elements $w_{ii}$ as the sum of net weights of all connections to movable module pin $i$.
The off-diagonal elements $w_{ij}$ are the negative sum of net weights of connections between movable module pins $i$ and $j$, resulting in a symmetric matrix.
The vector $\boldsymbol{c}_x$ has entry $-c_i$ in the $i$th row, where $c_i$ is the weight of the connection between fixed pin $i$ and and movable pin $i$, most of which will be zero.
The SimPL paper offers a small example of this construction for a toy netlist \cite{SimPL}.
